The effects of ongoing context inference on motor learning

### 2021-10-22 adaptation/article/article.tex
5k words reached. All good

\documentclass[a4paper,doc,floatsintext,natbib]{apa6}
% \documentclass{article, a4paper}
\usepackage[font=small]{caption}
\usepackage{lscape}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{authblk}
\usepackage[utf8]{inputenc}
\usepackage{nameref}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{soul}
\usepackage{todonotes}

% Remember to start reftex-mode

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=black,
  urlcolor=black
  }

@@

\title{The effects of context inference on motor learning: savings, de-adaptation and spontaneous recovery}
\shorttitle{Context-inference-dependent motor learning}
\author[1,2]{Cuevas Rivera, Darío}
\author[1,2]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
\affiliation{~}

@@

\section{Introduction}
It has been shown that humans can learn multiple tasks during the course of a single experiment, as well as learn to adapt to often-opposing modifications to the dynamics of movements \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}, although interference between tasks can slow down learning \citep{Brashers-Krug_Consolidation_1996,Sing_Reduction_2010}. Additionally, humans have been shown to dynamically switch between different learned adaptation \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.

In order to study motor adaptation, motor error is defined in experiments as the deviation between the observed movements of a participant and a baseline movement. For example, in experiments with a mechanical arm \citep[e.g.][]{Gandolfo_Motor_1996}, a correlation coefficient can be defined between the movements when the mechanical arm exerts a force and the movements when it does not; the higher the correlation coefficient, the lower the error. Other examples include measuring the force profile in error-clamp trials \citep{Smith_Interacting_2006}, endpoint errors in visuomotor rotations \citep{Kim_Neural_2015} and saccade adaptation \citep{Catz_Cerebellardependent_2008}.

By introducing blocks of trials in which a manipulation is present (e.g. a mechanical arm exerts a force), experimenters are able to observe the dynamics of adaptation through the lens of motor error. Across many different motor adaptation experiments, now well-established phenomena have been observed, for example: (i) savings (the ability to recall previously-learned skills), (ii) quick de-adaptation (the ability to return to unmodified dynamics), (iii) anterograde interference (the interference in motor learning between opposing manipulations in dynamics) and (iv) spontaneous recovery (during trials where errors are forced to be zero, the display of behavior consistent with the previously-learned dynamics) \citep[e.g.][]{Davidson_Scaling_2004}.

To explain these phenomena, a number of computational models have been introduced. The least complex of these models are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}. Bayesian accountshave also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between internal models of adaptation via predictions generated by these internal models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.

@@

In this work, we show that these unexplained phenomena, alongside savings, de-adaptation, anterograde interference and spontaneous recovery, can be explained in terms of the concept of context, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}. Intuitively, in mechanical arm experiments, one can think of one context in which the arm exerts no force and one in which it does. By inferring the current context, a participant can know how to counteract the mechanical arm and succeed in a reaching movement. In some experiments, the context is clearly inferable via sensory cues \cite[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Ethier_Spontaneous_2008}. In many other experiments, the context is not directly observable, and context inference is probabilistic and takes the form of an evidence-accumulating process across multiple trials. In this work, we propose that the behavioral consequences of this slow process of evidence accumulation can be observed in behavior and can be used to explain the phenomena outlined above.

To this end, we present a computational model for motor adaptation inspired by models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}. We build on these models by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference. In our model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions. To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment (e.g. whether the task was successfully completed).

@@

\section{Results}
In this work, we present a motor adaptation model in which the agent (e.g. a human learning a motor task) updates an existing model of the environment based on error signals produced by the need of adaptation. Several models exist for such error-based motor adaptation \cite[e.g.][]{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019}. We expand on these existing models by adding an explicit component of context inference. As we will show, this context inference guides the selection of the adequate internal model, its updating and the sampling of actions.

In what follows, we show that the proposed model explains several experimental phenomena found in previous studies. We simulated representative experiments from these studies to illustrate how our model explains these experimental findings using the dynamics of context inference. We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.

@@

\subsection{Modeling context-dependent adaptation}
The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these component happen in this order, and each component informs the ones that follow: motor adaptation is informed by context inference, and action selection is informed directly by both motor adaptation and by context inference.

Central to the model is the concept of context. In our model, contexts are defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the variables of the environment. Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to infer the context. For example, when lifting an object of unknown weight, an agent might have learned one context for heavy objects and one for light objects. When observing an object to be lifted, the agent can use its size and texture to estimate the weight of the object, which in turn allows the agent to infer the appropriate context and, with it, decide how to lift the object. These contextual cues (size and texture) do not necessarily identify the context with certainty (e.g. it could be a movie prop that looks like a rock) and integration over time might be necessary. All the variables of the model are fully described in the Methods section.

@@

\subsubsection{Motor adaptation}
Motor adaptation refers to the ability of humans and other animals to adapt their motor commands based on performance errors. For example, we can learn to perform a reaching movement when underwater, where the relatively high viscosity of water means that the motor commands we have learned all our lives no longer produce the desired effect.

It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. Here, we use this principle and chose to make use of an exact Bayesian learner. To adapt to altered dynamics (e.g. a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model. These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation. With this, a posterior estimate on $\phi$, $q(\phi)$, is calculated after each trial.

\subsubsection{Context inference}
In our model, the idea of context is quite general (see \fref{fig:model}A). The context comprises the relevant elements of the environment (e.g. the strength and direction of the wind when walking), of the task at hand (e.g. immediate and long-term goals and reward structure) and task-related information internal to the decision-making agent (e.g. forward model, decision-making algorithms, motor adaptation equations, expected observations). Most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the internal models: given a past action, the different internal models (e.g. two forward models $f_B$ and $f_H$) make different predictions for the outcome of that action; the model that best predicts that action is deemed to be the ``correct'' model, i.e. the model that best represents the context. In addition, \cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.

Following these ideas, our model performs context inference that integrates several sources of information, including sensory cues, proprioceptive information, outcome prediction errors and reward prediction errors. As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.

This component of the model can be used to include contextual sensory information from many sources. In experimental settings, for example, the contextual sensory information can take the form of a visual cue \citep[e.g.][]{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep[e.g.][]{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or even the start of a new block of trials \citep{Ethier_Spontaneous_2008}. All these sources of information, alongside the outcomes of previous actions and the prior beliefs over context $p(\vartheta)$ are integrated to infer the identity of a context.

@@

\subsection{Experimental results}
In this section, we present experimentally-observed phenomena in four sections, and show that context inference, as done in our model, provides a unifying explanation for all of them. In the first section, we discuss switches between contexts, and how slow context inference affects these switches. In the second section, we focus on the effects of volatility in the environment on motor adaptation. In the third section, we discuss the effects of slow context inference on action selection. Finally, in the fourth section we discuss context inference during so-called error-clamp trials, and its effect on behavior. For each of the four sections, we selected one or two studies which are representative of the phenomenon being discussed; a more extensive review of the literature can be found in the Discussion.

For clarity, we first introduce necessary terminology that is typically used in experimental studies and will be useful for linking experimental to computational observations. As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force (i.e. a force that pushes the participant's hand in the direction perpendicular to movement). Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Context A can be defined as that in which the arm exerts the clockwise curl force and context B counter-clockwise. Additionally, abusing notation, the usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions. Studies often talk of contexts such as A and A/2, which means that both have the same direction of the adaptation (e.g. clockwise), but the second one has half the magnitude. Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter E.

With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of O-A-B-E, or O-A-(-A)-E, which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E).

In some experiments, adaptation blocks are repeated throughout the experiments. For example, \cite{Oh_Minimizing_2019} used an O-A-O-A-O-A-O-A sequence. To differentiate the repeating blocks, the authors named them $O_i$ and $A_i$, where $i = 1, 2, ...$. Their experiment can be described as $O_1$-$A_1$-$O_2$-.... We adopt this naming convention.

We use the term "learned adaptation" and similar to refer to motor outputs (e.g. reaching movements) that have already been adapted by the participant to a change in the dynamics of the environment. For example, the learned adaptation during a task with a curl force refers to the movements of the participant that already counteract the force exerted by the mechanical arm.

\subsubsection{Savings and slow/fast switching}
The term 'savings' refer to the ability of humans and other animals to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in human participants \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block.

In this section, we discuss savings, and the related phenomenon of de-adaptation, in terms of switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}. Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not.

@@

Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, in curl-force-field experiments with mechanical arms, the force feedback provided by the mechanical arm always gives participants a sense for the current context, although error-clamp trials are difficult to distinguish from adaptation trials based on force feedback alone. In fact, the mere act of holding the handle informs participants that something is different from baseline. Proprioceptive signals can also provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change, as long as the motor error is perceivably higher than the expected trial-by-trial variation in performance, and even an unusually long pause between two trials could indicate to participants that a new block with a potentially different context could have started \cite{Ethier_Spontaneous_2008}.

To our knowledge, there are no experiments in which no contextual information is available. However, many experiments \citep[e.g.][]{Huang_Persistence_2009,Brennan_Decay_2015,Smith_Interacting_2006} contain blocks of error-clamp trials in which participants are not provided with contextual information. These blocks could be considered uncued, however, as we discuss in the following sections, contextual information is nevertheless available to participants. In this section, we focus on cued and partially-cued experiments and leave error-clamp trials to the following sections.

To show how adaptation changes depending on the amount of contextual information, we selected three representative experiments from two studies: \cite{Kim_Neural_2015} and \cite{Oh_Minimizing_2019} which differ in the amount of contextual information. \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, the identity of the current context was displayed as a visual cue (colored light), making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e. when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.

@@

Critically, in \fref{fig:oh-2019}A it can be seen until trial 250 and especially for the first four context switches,  that participants did not completely adapt to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks. Switching between opposite adaptations was nevertheless immediate in simulations and experimental data. After around 500 trials, participants had adapted completely, as seen by the blue line reaching the heights of the black line, and switches continued to be immediate.

These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches. To expand on these results, we now turn to performance feedback and its effects on switching behavior.

\cite{Oh_Minimizing_2019} performed partially-cued experiments similar to \cite{Kim_Neural_2015}. The authors performed two experiments that differed in the size of the adaptation: in the first experiment, the angle of rotation was 20 degrees, while in the second experiment it was only 10 degrees. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, respectively, alongside simulations with our model. Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A_1$. From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (true rotation). In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation (blue line) lagging behind the switches in the real context (black line), and slowly catching up. As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C). This difference is explained by our model in terms of the size of the adaptation: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism in our model.

\subsubsection{Uncertainty over contexts affects action selection}
As with learning, our model selects actions (motor commands) based on context inference. If the identity of the current context is known, the forward model for this context is used to select the current action. However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eq:action-selection}).

Experimental evidence supporting this view can be found in experiments with context switching. For example, \cite{Davidson_Scaling_2004} reported an curl-force experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A). After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A. The authors interpreted this as evidence that switching between adaptations happens more quickly if it is in the same direction as the current adaptation (e.g. both counter-clockwise), and more slowly if they are in the opposite direction (e.g. clockwise to counter-clockwise).

Our model offers a different explanation of the observed results, based on context inference: the asymmetry is due to the existence of the baseline context. As discussed above, action selection in our model is done via a weighted average (see \eref{eq:action-selection}). In experiments without cued contexts, the baseline context O and its associated forward model $m_O$, have a non-zero probability $p(\zeta | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $m_O$ becomes more likely (given that $m_{3A}$ has been ruled out). Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$.

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}, shown in \fref{fig:davidson-2004}B. Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data. However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.

@@

\subsubsection{Action selection in error-clamp blocks}
During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Spontaneous recovery, during which participants' behavior is consistent with a previously-encountered context (e.g. in an O-A-E experiment, consistent with A); this phase, when present, is seen during the early trials of the E block. (2) A slow return to baseline (i.e. no observable adaptation, consistent with the baseline context), which can last as long as hundreds of trials \citep{Brennan_Decay_2015}. However, the direction of the spontaneous recovery, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation seem to vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.

In this section, we show how our model can explain these different parameters of behavior by changing the way contextual cues mislead participants' context inference, which in turn influences action selection.

>>>>

\documentclass[a4paper,doc,floatsintext,natbib]{article}
% \documentclass{article}
\usepackage[font=small]{caption}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage[utf8]{inputenc}
\usepackage{nameref}
\usepackage{todonotes}
\usepackage{authblk}
\usepackage[nomarkers,figuresonly]{endfloat}

% Remember to start reftex-mode
\bibliographystyle{ieeetr}

@@

\title{The effects of ongoing context inference on motor learning}
% \shorttitle{Context-inference-dependent motor learning}
\author[1,2]{Cuevas Rivera, Darío}
\author[1,2]{Kiebel, Stefan J.}
\affil[1]{Chair of Neuroimaging, Faculty of Psychology. Technische Universität Dresden, 01062 Dresden, Germany.}
\affil[2]{Centre for Tactile Internet with Human-in-the-Loop (CeTI)}
% \affiliation{~}

@@

\section{Introduction}
It has been shown that humans can adapt motor commands to counteract changes in the dynamics of the environment and their own bodies, such as performing reaching movements with a weight attached to the wrist. This is known as motor adaptation. Moreover, human participants have been shown to adapt to different, even opposing, such changes during the course of a single experiment \citep{Gandolfo_Motor_1996,Shadmehr_Functional_1997}. Additionally, humans have been shown to dynamically switch between different learned adaptations \citep{Davidson_Scaling_2004,Ethier_Spontaneous_2008,Lee_Dual_2009}.

By introducing blocks of trials in which body dynamics are altered (e.g. a mechanical arm exerts a force on the participant's hand), experimenters are able to observe motor adaptation through the lens of motor error. Across many different motor adaptation experiments \citep[e.g.][]{Gandolfo_Motor_1996,Shadmehr_Adaptive_1994,Davidson_Scaling_2004}, well-established phenomena have been observed: (i) the ability to recall previously-learned skills, called savings; (ii) the ability to return to unmodified dynamics, termed de-adaptation; (iii) the interference in motor learning between opposing manipulations in dynamics, called anterograde interference; (iv) spontaneous display of behavior consistent with a previously-learned adaptation, during trials where errors are forced to be zero, called spontaneous recovery.

To explain these phenomena, a number of computational models have been introduced, which adapt their motor commands after observing motor errors. The most studied are linear learners \citep{Smith_Interacting_2006,Forano_Timescales_2020,Scheidt_Learning_2001}, but Bayesian accounts have also been presented, providing an alternative explanation for savings and quick de-adaptation in the form of switching between forward models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019}.

@@

In this work, we show that these unexplained phenomena can be explained in terms context inference, an idea that has recently been gaining traction in different fields \cite[e.g.][]{Sanders_Hippocampal_2020,Hunter_Contextsensitive_2021}. We build on models such as MOSAIC \citep{Wolpert_Multiple_1998} and Bayesian switching models \citep{Kording_Bayesian_2004,Oh_Minimizing_2019} by expanding the concept of context and introducing an explicit mechanism for inferring the context based on Bayesian inference. In the proposed model, the context encompasses all the information relevant to making decisions, both environmental and internal to the agent making these decisions. To infer the current context, our model integrates several sources of information, including the predictions made by its internal forward models, an estimate of the reliability of these predictions, direct sensory information (such as visual contextual cues), proprioceptive information, and feedback provided by the environment.

@@

\section{Results}
We present a motor adaptation model in which an existing model of the environment is updated based on error signals produced by the need of adaptation. We expanded on existing models \citep{Wolpert_Multiple_1998,Kording_Bayesian_2004,Oh_Minimizing_2019} by adding an explicit component of context inference. As we will show, context inference guides the selection of the adequate forward model, its updating and the sampling of actions.

We simulated representative experiments from a number of experimental studies on motor adaptation to illustrate how our model explains these experimental findings using the dynamics of context inference. We will present these simulations alongside the experimental results from the representative studies and discuss in detail how context inference explains the experimental phenomena.

@@

\subsection{Modeling context-dependent adaptation}
The model we present has three main components: (1) context inference, (2) motor adaptation and (3) action selection. The processes defined by these component occur in this order, and each component informs the ones that follow.

Central to the model is the concept of context, defined in terms of the task to be performed (see \fref{fig:model}A), the variables of the environment that are relevant to perform the task, the forward models used by the decision-making agent to perform the task, and the update mechanisms necessary to adapt these forward models to the changing environment. Together, these elements allow the agent to make predictions on future observations when this context is active, and these predictions are used to infer the context. For example, when lifting an object of unknown weight, an agent might have learned one context for heavy objects and one for light objects. When observing an object to be lifted, the agent can use its size and texture to estimate the weight of the object, which in turn allows the agent to infer the appropriate context and, with it, decide how to lift the object.

@@

\subsubsection{Motor adaptation}
It is widely accepted that in order to adapt motor commands to a novel environment, we create and update internal forward models of the outcomes of control signals \citep{Wolpert_Multiple_1998}. We used this principle and chose to make use of an exact Bayesian learner. To adapt to altered dynamics (e.g. a mechanical arm exerting a force on a participant's hand), our model updates the parameters $\phi$ (see Methods) of a context-specific forward model. These updates are done according to \eref{eqn:motor-adaptation}, using the difference between predictions made with the current parameters of the forward model, and the current observation. With this, a posterior estimate on $\phi$, $q(\phi)$ in \fref{fig:model}B and C, is calculated after each trial.

\subsubsection{Context inference}
To infer the adequate forward model to deploy, most previous models \cite[e.g.][]{Wolpert_Multiple_1998,Imamizu_Neural_2008,Oh_Minimizing_2019} have focused on the elements of the context which are directly related to the forward models: the model that best predicts the current observation given the past is deemed to be the ``correct'' model, i.e. the model that best represents the current context. \cite{Imamizu_Explicit_2007} expanded on this idea by showing the effects of visual cues on model switching.

Following these ideas, our model performs context inference that integrates several sources of information, including those typical for experimental studies such as visual cues \citep{Lee_Dual_2009,Kim_Neural_2015}, the place where the task must be carried out \citep{Forano_Timescales_2020,Shadmehr_Adaptive_1994} or the start of a new block of trials \citep{Ethier_Spontaneous_2008}, in addition to proprioceptive information, outcome prediction errors and reward prediction errors. As with motor adaptation, the context is inferred with exact Bayesian inference, with the updates given by \eref{eqn:estimated-context}.

@@

\subsection{Experimental results}
In this section, we present experimentally-observed phenomena in three sections, and show that the dynamics of context inference provide a unifying explanation for all of them. In the first section, we discuss switches between contexts, and how slow context inference affects these switches. In the second section, we focus on the effects of volatility in the environment on motor adaptation. Finally, in the third section we discuss context inference during so-called error-clamp trials, and its effect on behavior. For each of the three sections, we selected one or two studies which are representative of the phenomenon being discussed.

For clarity, we first introduce necessary terminology that is typically used in experimental studies. As an example, we will use a typical motor adaptation task in which participants have to make reaching movements while holding the handle of a mechanical arm that exerts a curl force on the participant's hand. Depending on the trial, the mechanical arm might exert a curl force in a clockwise or counter-clockwise direction, or no force at all. Let us define the baseline context O as that in which the mechanical arm exerts no force. Contexts A and B can be defined as those with clockwise or counter-clockwise, respectively. Abusing notation, a usual statement is that $B = (-A)$, as the forces have the same magnitude but point in opposite directions. Similarly, one can define context A/2, with the same direction of adaptation as A, but half the magnitude. Finally, many experiments include a block of error-clamp trials at the end of the experiment, in which the mechanical arm forces the participant to make straight-line movements; we represent these with the letter E.

With this terminology, a typical experiment \cite[e.g.][]{Ethier_Spontaneous_2008} would have a block structure of O-A-B-E, or O-A-(-A)-E, which means that the participant goes through a block of trials with no external force applied (O), a number of trials with a clockwise curl force (A), a block with counter-clockwise forces (B), and finally a block with error-clamp trials (E). With repeated contexts \citep[e.g.][]{Oh_Minimizing_2019}, an experiment can be described as $O_1$-$A_1$-$O_2$-\ldots .

\subsubsection{Savings and slow/fast switching}
The term 'savings' refer to the ability to remember a previously-learned adaptation and apply it without having to re-learn it. Savings is almost universally observed in humans \citep{Brashers-Krug_Consolidation_1996,Shadmehr_Functional_1997,Medina_Mechanism_2001,Smith_Interacting_2006,Zarahn_Explaining_2008}. In an O-A-O-A experiment, for example, savings would express themselves in the second A block in the form of a much higher adaptation rate higher than that observed during the first A block. The related concept of quick de-adaptation occurs in A-O transitions, where participants switch back to baseline without having to re-learn it.

In this section, we discuss savings and de-adaptation in terms of switching between forward models. This has been modeled before \citep[e.g.][]{Wolpert_Multiple_1998,Oh_Minimizing_2019,Imamizu_Explicit_2007}. Here we show in addition that expanding forward model switching into context inference explains facets of savings that these previous studies did not, namely that savings are not immediate, but a relatively fast process that reflects context inference.

@@

We selected three representative experiments from two studies \citep{Kim_Neural_2015,Oh_Minimizing_2019} which differ in the amount of contextual information available to participants. \cite{Kim_Neural_2015} performed a cued-context visuomotor rotation experiment with three contexts with different rotation: no rotation, 40 degrees and -40 degrees. Participants performed shooting movements in blocks of trials with the same rotation. Importantly, a colored light identified the current context, making this a cued-context experiment. Consistent with our model, the authors found that switching was immediate and accurate. In \fref{fig:oh-2019}A we show the results of simulations with our model using the parameters of the task, as well as the experimental results from \cite{Kim_Neural_2015}. The correspondence between simulations and experimental results can be seen in the switches between contexts, i.e. when the solid black line (representing the true context) switches between 40, 0 and -40; in both panels (left and right), the blue line, representing the agent's adaptation, quickly follows the switches.

@@

Critically, in \fref{fig:oh-2019}A it can be seen for the first four context switches (until about trial 250) that participants had not yet completely adapted to the rotation, as evidenced by the blue line not reaching as low or high as the black line during the first blocks. Switching between opposite adaptations was nevertheless immediate in simulations and experimental data. After around 500 trials, participants had adapted completely, as seen by the blue line reaching the heights of the black line, and switches continued to be immediate. These results are similar to those shown by \cite{Imamizu_Explicit_2007}, where sensory cues of varying reliability effected immediate or slow contextual switches. To expand on these results, we now turn to performance feedback and its effects on switching behavior.

\cite{Oh_Minimizing_2019} performed two partially-cued experiments with a visual rotation of 20 and 10 degrees, respectively. The results of their experiments can be seen in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, alongside simulations with our model. Participants in the first experiment (\fref{fig:oh-2019}B) first learned the adaptation in $A_1$. From the $A_1 - O_2$ transition onward, participants showed immediate switching (with a one-trial lag) between A and O (both ways), which can be seen in terms of the blue line (adaptation) closely following the switches in the black line (true rotation). In the second experiment (\fref{fig:oh-2019}C), switching happens more slowly, with adaptation lagging behind the switches in the real context, and slowly catching up. As can be seen in the left panels in \fref{fig:oh-2019}B and \ref{fig:oh-2019}C, the same model parametrization produces fast, accurate switches when the adaptation is large (B), and slow, noisy switches when it is low (C). This difference is explained by our model in terms of the size of the adaptation: as the adaptation is smaller (10 degrees), it becomes more difficult to distinguish errors made by incorrectly inferring the context from the noise due to trial-to-trial variation in motor output. Because of this, the model requires more evidence (i.e. more trials) to infer a switch in contexts.

Note that the results from our simulations, from \cite{Oh_Minimizing_2019} and from \cite{Kim_Neural_2015} include both savings (O-A transitions) and de-adaptation (A-O transitions), both of which display the same characteristics and are explained by the same mechanism.

\subsubsection{Uncertainty over contexts affects action selection}
As with learning, our model selects actions (motor commands) based on context inference. If the identity of the current context is known, the forward model for this context is used to select the current action. However, if uncertainty over the context exists, the selected action is influenced by all the possible contexts, with a weight directly related to how likely each one of those contexts is (see \eref{eqn:dist-comm}).

Experimental evidence supporting this view can be found in experiments with context switching. For example, \cite{Davidson_Scaling_2004} reported a curl-force experiment in which participants had to switch from 3A to A in one group, with a block sequence A-3A-A-3A, and from -A to A in another group, with a block sequence A-(-A)-A-(-A). After A and 3A (or -A in the other group) had been learned in the first two blocks, the authors found that the switch from 3A to A was faster than that from -A to A. The authors interpreted this as evidence that switching between adaptations happens more quickly if it is in the same direction as the current adaptation (e.g. both counter-clockwise), and more slowly if they are in the opposite direction (e.g. clockwise to counter-clockwise).

In our context-inference account, the asymmetry is caused by the baseline context, which have a non-zero probability $p(\zeta_O | s_t ...)$. When a new block of trials starts (e.g. in the transition from 3A to A), a switch is inferred by the model (given feedback after the first trial) and $\zeta_O$ becomes more likely (given that $\zeta_{3A}$ has been ruled out). Therefore, in these first trials, action selection has a component guided by the baseline model, in which no extra compensatory force is applied, effectively ``pulling'' adaptation towards zero (no compensatory force). In the first group, this initial pull towards zero accelerates the transition towards A because $3A > A > 0$, but in the second group, it slows down the switch because $A > 0 > -A$.

We simulated data with our model fitting the experimental structure in \cite{Davidson_Scaling_2004}. We show the results in \fref{fig:davidson-2004}, alongside the experimental results from \cite{Davidson_Scaling_2004}. It can be seen that the agent exposed to the -A context shifts back to A more slowly than the one exposed to 3A, qualitatively reproducing the data from \cite{Davidson_Scaling_2004}. Note that in our simulations, the difference between the groups, as well as the variability within each group, is smaller than in the experimental data. However, the effect can be reproduced with any number of simulated runs (each one representing one participant), pointing towards a reliable effect.

@@

\subsubsection{Action selection in error-clamp blocks}
During error-clamp blocks at the end of block sequences, participants' behavior can be divided in two phases: (1) Spontaneous recovery, during which participants' behavior is consistent with a previously-encountered context (e.g. in an O-A-E experiment, consistent with A); this phase, when present, is seen during the early trials of the E block. (2) A slow return to baseline, which can last as long as hundreds of trials \citep{Brennan_Decay_2015}. However, the direction of the spontaneous recovery, its duration, the delay before it is observed, the speed of the return to baseline and the final asymptote of the adaptation vary greatly depending on the experiment \citep{Brennan_Decay_2015,Vaswani_Decay_2013,Smith_Interacting_2006,Shmuelof_Overcoming_2012}.

Because context inference integrates information from different sources, many experiments in which no intentional, overt contextual cues are available indeed contain contextual information that the participant can use to infer the context. For example, proprioceptive signals  provide contextual information \citep{Dizio_Motor_1995,Shadmehr_Adaptive_1994}. The sudden appearance of motor errors can itself be a cue for contextual change \citep{Herzfeld_memory_2014} and even a pause between two trials could suggest a change in context \cite{Ethier_Spontaneous_2008}.

In this section, we show how context inference can explain these different parameters of behavior by changing the way contextual cues mislead participants' context inference, which in turn influences action selection.
